---
title: "Madison_Wisconsin_Data"
author: "Maddy Enda"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

#loading in libraries and data
```{r}
library(tidyverse)
library(here)
library(janitor)
library(patchwork)

ice_data <- read_csv(here("data","Ice_duration_data.csv"), na = c("-999", "")) %>% 
  clean_names() 

meteor_data <- read_csv(here("data","Madison_Daily_Meteorological_Data.csv"), na = c("-999", "")) %>% 
  clean_names() 
```


#creating summary data for ice cover
```{r}

#visualizing summary data for ice cover
```{r}
ggplot(ice_data, aes(x= lakeid, y= ice_duration)) + 
geom_boxplot() +
  stat_summary(fun.y= mean, geom= "point", fill= "red", shape= 23, size= 2)

```

#Finding mean ice duration
```{r}
ice_data_summary <- ice_data %>%
  group_by(year4) %>%
  summarise(mean_ice_duration= mean(ice_duration, na.rm = TRUE)) 
```

#Visualizing ice data by year
```{r}
ggplot(ice_data_summary, aes(x= year4, y= mean_ice_duration)) + geom_line()

#The overall trend appears to be that the average duration of time that ice coverage remains is decreasing over time. 
```

#Finding Mean Air temp
```{r}
madison_air_summary <- meteor_data %>%
  filter(month %in% c(1,2, 12)) %>%
  select(ave_air_temp_adjusted, year4, month) %>%
  group_by(year4) %>%
  summarize( mean_air_temp= mean(ave_air_temp_adjusted, na.rm= TRUE))

```

#Visualizing Mean air temp
```{r}
ggplot(madison_air_summary, aes(x= year4, y= mean_air_temp)) +
  geom_line()

#The overall trend appears to be that the average air temperature throughout the winter months in Madison is increasing over time
```

#Joining the data together
```{r}
Madison_ice_air_data <- ice_data_summary %>%
left_join(madison_air_summary)
```

#Visualizing data
```{r}
a <- ggplot(data= Madison_ice_air_data, aes(x= year4, y= mean_air_temp)) +
  geom_line(color= "aquamarine") +
  theme_dark()

i <- ggplot(data= Madison_ice_air_data, aes(x= year4, y= mean_ice_duration)) + 
  geom_line(color= "white") +
  theme_dark()

madison_air_ice_graph <-(a | i) 
```

#editing the graph
```{r}
madison_air_ice_graph %>%
 #hi there 
```


